Probabilistic Symbolic Execution

### 2021-06-30 shared/Paper-Draft/symbolic_execution.tex
WIP : SymbEx Section (B)

%% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
%%\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
%% For double-blind review submission, w/ CCS and ACM Reference
%\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true}
%% For single-blind review submission, w/o CCS and ACM Reference (max submission space)
%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
%% For single-blind review submission, w/ CCS and ACM Reference
%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true}
%% For final camera-ready submission, w/ required CCS and ACM Reference
\documentclass[acmsmall,review]{acmart}\settopmatter{}

%% Journal information
%% Supplied to authors by publisher for camera-ready submission;
%% use defaults for review submission.
\acmJournal{PACMPL}
\acmVolume{1}
\acmNumber{POPL} % CONF = POPL or ICFP or OOPSLA
\acmArticle{1}
\acmYear{2022}
\acmMonth{1}
\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
\startPage{1}

%% Copyright information
%% Supplied to authors (based on authors' rights management selection;
%% see authors.acm.org) by publisher for camera-ready submission;
%% use 'none' for review submission.
\setcopyright{none}
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\copyrightyear{2018}           %% If different from \acmYear

%% Bibliography style
\bibliographystyle{ACM-Reference-Format}
%% Citation style
%% Note: author/year citations are required for papers published as an
%% issue of PACMPL.
\citestyle{acmauthoryear}   %% For author/year citations

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Note: Authors migrating a paper from PACMPL format to traditional
%% SIGPLAN proceedings format must update the '\documentclass' and
%% topmatter commands above; see 'acmart-sigplanproc-template.tex'.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% Some recommended packages.
\usepackage{booktabs}   %% For formal tables:
%% http://ctan.org/pkg/booktabs
\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
%% http://ctan.org/pkg/subcaption
\usepackage{mathtools}
\usepackage[noend]{algpseudocode}
\usepackage{algorithm}
\usepackage{forest}
\usepackage{csquotes}
\usepackage{minted}
\usepackage{multicol}
\usepackage{xcolor}

%%% Colors %%%
\definecolor{darkgreen}{RGB}{0,128,0}

%%% Macros %%%

\DeclarePairedDelimiter{\abs}{\lvert}{\rvert}
\newcommand{\DeclareAutoPairedDelimiter}[3]{%
	\expandafter\DeclarePairedDelimiter\csname Auto\string#1\endcsname{#2}{#3}%
	\begingroup\edef\x{\endgroup
		\noexpand\DeclareRobustCommand{\noexpand#1}{%
			\expandafter\noexpand\csname Auto\string#1\endcsname*}}%
	\x}
\DeclareAutoPairedDelimiter{\deno}{\llbracket}{\rrbracket}

\newcommand*\Let[2]{\State #1 $\gets$ #2}
\newcommand*\cLet[3]{\State\ColorLine{#3} #1 $\gets$ #2}
\newcommand*\dom[1]{\mathrm{dom}\left( #1 \right)}
% For the Case statements
% Switch Statement Commands
% New definitions
\algnewcommand\algorithmicswitch{\textbf{switch}}
\algnewcommand\algorithmiccase{\textbf{case}}
\algnewcommand\algorithmicassert{\texttt{assert}}
\algnewcommand\Assert[1]{\State \algorithmicassert(#1)}%
% New "environments"
\algdef{SE}[SWITCH]{Switch}{EndSwitch}[1]{\algorithmicswitch\ #1\ \algorithmicdo}{\algorithmicend\ \algorithmicswitch}%
\algdef{SE}[CASE]{Case}{EndCase}[1]{\algorithmiccase\ #1}{\algorithmicend\ \algorithmiccase}%
\algtext*{EndSwitch}%
\algtext*{EndCase}%

% Highlight lines in algorithm environment
\makeatletter
% code borrowed from Andrew Stacey; See
% https://tex.stackexchange.com/a/50054/3954
\tikzset{%
	remember picture with id/.style={%
		remember picture,
		overlay,
		save picture id=#1,
	},
	save picture id/.code={%
		\edef\pgf@temp{#1}%
		\immediate\write\pgfutil@auxout{%
			\noexpand\savepointas{\pgf@temp}{\pgfpictureid}}%
	},
	if picture id/.code args={#1#2#3}{%
		\@ifundefined{save@pt@#1}{%
			\pgfkeysalso{#3}%
		}{
			\pgfkeysalso{#2}%
		}
	}
}

\def\savepointas#1#2{%
	\expandafter\gdef\csname save@pt@#1\endcsname{#2}%
}

\def\tmk@labeldef#1,#2\@nil{%
	\def\tmk@label{#1}%
	\def\tmk@def{#2}%
}

\tikzdeclarecoordinatesystem{pic}{%
	\pgfutil@in@,{#1}%
	\ifpgfutil@in@%
	\tmk@labeldef#1\@nil
	\else
	\tmk@labeldef#1,(0pt,0pt)\@nil
	\fi
	\@ifundefined{save@pt@\tmk@label}{%
		\tikz@scan@one@point\pgfutil@firstofone\tmk@def
	}{%
		\pgfsys@getposition{\csname save@pt@\tmk@label\endcsname}\save@orig@pic%
		\pgfsys@getposition{\pgfpictureid}\save@this@pic%
		\pgf@process{\pgfpointorigin\save@this@pic}%
		\pgf@xa=\pgf@x
		\pgf@ya=\pgf@y
		\pgf@process{\pgfpointorigin\save@orig@pic}%
		\advance\pgf@x by -\pgf@xa
		\advance\pgf@y by -\pgf@ya
	}%
}

\makeatother
% end of Andrew's code

% main command to draw the colored background
\newcounter{mymark}
\newcommand\ColorLine[1]{%
	\stepcounter{mymark}%
	\tikz[remember picture with id=mark-\themymark,overlay] {;}%
	\begin{tikzpicture}[remember picture,overlay]%
		\filldraw[#1]%
		let \p1=(pic cs:mark-\themymark), 
		\p2=(current page.east)  in 
		([xshift=-0.3em,yshift=-0.7ex]0,\y1)  rectangle ++([xshift=-1.2cm]\x2,\baselineskip);
	\end{tikzpicture}%
}%

% colored loops and declarations
\makeatletter
\algnewcommand\cRequire[1]{\item[\ColorLine{#1}\algorithmicrequire]}%
\algnewcommand\cEnsure[1]{\item[\ColorLine{#1}\algorithmicensure]}%
\algnewcommand\cState[1]{\State\ColorLine{#1}}%
\algnewcommand\cStatex[1]{\Statex\ColorLine{#1}}%
\algnewcommand\cComment[1]{\Comment\ColorLine{#1}}%

\algdef{SE}[WHILE]{cWhile}{ENDWHILE}%
[2]{\ColorLine{#2}\algorithmicwhile\ #1\ \algorithmicdo}%
{\algorithmicend\ \algorithmicwhile}%
\algdef{SE}[FOR]{cFor}{ENDFOR}%
[2]{\ColorLine{#2}\algorithmicfor\ #1\ \algorithmicdo}%
{\algorithmicend\ \algorithmicfor}%
\algdef{S}[FOR]{cForAll}%
[2]{\ColorLine{#2}\algorithmicforall\ #1\ \algorithmicdo}%
\algdef{SE}[LOOP]{cLoop}{ENDLOOP}%
[1]{\ColorLine{#1}\algorithmicloop}%
{\algorithmicend\ \algorithmicloop}%
\algdef{SE}[REPEAT]{cRepeat}{UNTIL}%
[1]{\ColorLine{#1}\algorithmicrepeat}%
[1]{\algorithmicuntil\ #1}%
\algdef{SE}[IF]{cIf}{ENDIF}%
[2]{\ColorLine{#2}\algorithmicif\ #1\ \algorithmicthen}%
{\algorithmicend\ \algorithmicif}%
\algdef{C}[IF]{IF}{cElsif}%
[2]{\ColorLine{#2}\algorithmicelse\ \algorithmicif\ #1\ \algorithmicthen}%
\algdef{Ce}[ELSE]{IF}{cElse}{ENDIF}%
[1]{\ColorLine{#1}\algorithmicelse}%
\algdef{SE}[SWITCH]{cSwitch}{EndSwitch}%
[2]{\ColorLine{#2}\algorithmicswitch\ #1\ \algorithmicdo}%
{\algorithmicend\ \algorithmicswitch}%
\algdef{SE}[CASE]{cCase}{EndCase}%
[2]{\ColorLine{#2}\algorithmiccase\ #1}%
{\algorithmicend\ \algorithmiccase}%
\makeatother

\begin{document}

	%% Title information
	\title[PSE]{Probabilistic Symbolic Execution}         %% [Short Title] is optional;
	%% when present, will be used in
	%% header instead of Full Title.
	%\titlenote{with title note}             %% \titlenote is optional;
	%% can be repeated if necessary;
	%% contents suppressed with 'anonymous'
	%\subtitle{Subtitle}                     %% \subtitle is optional
	%\subtitlenote{with subtitle note}       %% \subtitlenote is optional;
	%% can be repeated if necessary;
	%% contents suppressed with 'anonymous'

	%% Author information
	%% Contents and number of authors suppressed with 'anonymous'.
	%% Each author should be introduced by \author, followed by
	%% \authornote (optional), \orcid (optional), \affiliation, and
	%% \email.
	%% An author may have multiple affiliations and/or emails; repeat the
	%% appropriate command.
	%% Many elements are not rendered, but should be provided for metadata
	%% extraction tools.

	%% Author with single affiliation.
	\author{First1 Last1}
	\authornote{with author1 note}          %% \authornote is optional;
	%% can be repeated if necessary
	\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
	\affiliation{
		\position{Position1}
		\department{Department1}              %% \department is recommended
		\institution{Institution1}            %% \institution is required
		\streetaddress{Street1 Address1}
		\city{City1}
		\state{State1}
		\postcode{Post-Code1}
		\country{Country1}                    %% \country is recommended
	}
	\email{first1.last1@inst1.edu}          %% \email is recommended

	%% Author with two affiliations and emails.
	\author{First2 Last2}
	\authornote{with author2 note}          %% \authornote is optional;
	%% can be repeated if necessary
	\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
	\affiliation{
		\position{Position2a}
		\department{Department2a}             %% \department is recommended
		\institution{Institution2a}           %% \institution is required
		\streetaddress{Street2a Address2a}
		\city{City2a}
		\state{State2a}
		\postcode{Post-Code2a}
		\country{Country2a}                   %% \country is recommended
	}
	\email{first2.last2@inst2a.com}         %% \email is recommended
	\affiliation{
		\position{Position2b}
		\department{Department2b}             %% \department is recommended
		\institution{Institution2b}           %% \institution is required
		\streetaddress{Street3b Address2b}
		\city{City2b}
		\state{State2b}
		\postcode{Post-Code2b}
		\country{Country2b}                   %% \country is recommended
	}
	\email{first2.last2@inst2b.org}         %% \email is recommended

	%% Abstract
	%% Note: \begin{abstract}...\end{abstract} environment must come
	%% before \maketitle command
	\begin{abstract}
		TODO: Write Abstract
	\end{abstract}

	%% 2012 ACM Computing Classification System (CSS) concepts
	%% Generate at 'http://dl.acm.org/ccs/ccs.cfm'.
	%% TODO: Get CSS Concepts
	\begin{CCSXML}
		<ccs2012>
		<concept>
		<concept_id>10011007.10011006.10011008</concept_id>
		<concept_desc>Software and its engineering~General programming languages</concept_desc>
		<concept_significance>500</concept_significance>
		</concept>
		<concept>
		<concept_id>10003456.10003457.10003521.10003525</concept_id>
		<concept_desc>Social and professional topics~History of programming languages</concept_desc>
		<concept_significance>300</concept_significance>
		</concept>
		</ccs2012>
	\end{CCSXML}

	\ccsdesc[500]{Software and its engineering~General programming languages}
	\ccsdesc[300]{Social and professional topics~History of programming languages}
	%% End of generated code

	%% Keywords
	%% comma separated list
	\keywords{keyword1, keyword2, keyword3}  %% \keywords are mandatory in final camera-ready submission

	%% \maketitle
	%% Note: \maketitle command must come after title commands, author
	%% commands, abstract environment, Computing Classification System
	%% environment and commands, and keywords command.
	\maketitle

	\section{Introduction}
	\label{sec:intro}

	One of the goals of symbolic execution is to identify whether a ``bad'' state (e.g. an \texttt{assert} failure) is reached, signifying a bug in the program.
	%
	In probabilistic programs, we are often more interested in either whether a ``bad'' state is reached too often, or, conversely, whether the ``good'' states are reached often enough.

	\subsection{Contributions}
	\label{sec:contrib}

	\section{Motivating Example}
	\label{sec:motiv_example}

	The Monty Hall problem~\cite{selvin1975} is a classic probability puzzle based on the American television show, \textit{Let's Make a Deal}, which showcases how subtle probabilistic reasoning can be.
	%
	The problem itself is simple:
	\begin{displayquote}
		You are a contestant on a gameshow and behind one of three doors there is a car and behind the others, goats.
		% 
		You pick a door and the host, who knows what is behind each of the doors, opens a different door, which has a goat.
		% 
		The host then offers you the choice to switch to the remaining door.
		% 
		Should you?
	\end{displayquote}
	While it may seem unintuitive, regardless of the contestant's original door choice, the contestant who always switches doors will win the car $\frac{2}{3}$ of the time, as opposed to a $\frac{1}{3}$ chance if the contestant sticks with their original choice.

	\begin{figure}\centering
		\inputminted[fontsize=\footnotesize,linenos]{c}{montyhall.c}
		\caption{C code for the Monty Hall Problem}
		\label{fig:montyhall}
	\end{figure}

	We can represent this problem as a probabilistic program, as shown in Fig.~\ref{fig:montyhall}, where $\mathtt{choice} \in [1,3]$ is the door which is originally chosen by the contestant and $\mathtt{door\_switch}$ is \texttt{true} if the contestant wants to switch doors when asked, and \texttt{false} otherwise.
	%
	If $\mathtt{door\_switch} = \mathtt{true}$, regardless of the value of \texttt{choice}, \texttt{monty\_hall} should return \texttt{true} (i.e. the car is won) $\frac{2}{3}$ of the time.
	%
	In general, the problem we aim to solve is: given a probabilistic program with discrete sampling statements, and a program property, how do we verify that the program satisfies the property?

	% One typical solution would be to run a simulation on random inputs and compute the probability of winning if $\mathtt{door\_switch} = \mathtt{true}$, and the probability of winning if $\mathtt{door\_switch} = \mathtt{false}$.
	% %
	% While for this program this would work as the input space is tiny, the resulting output distributions would be absent any formal guarantees and only be an approximation.
	% %
	% We instead develop a novel symbolic execution technique to provide a sound guarantee on the probability of certain events occurring.

	In symbolic execution, program inputs are replaced by \textit{symbolic} variables which can take on any value.
	%
	The program is then ``run'' on these symbolic variables and when a branch is reached execution proceeds along each of these two branches and the constraint is recorded in that path's, \textit{path condition}, which is represented by $\varphi$.
	%
	This yields an execution tree which shows us possible paths through the program.

	\begin{figure}
		\label{fig:montyhall_tree}
		\centering
		{\footnotesize
			\begin{forest}baseline,for tree=draw,
				[{$\mathtt{1}: c_0=\top$},align=center, base=bottom
				[{$\mathtt{4}: c_1 = \alpha = 1$},align=center, base=bottom
				[{$\mathtt{5}: c_2 = \beta$},align=center, base=bottom, node options={dotted,thick}, draw=red] %loss
				[{$\mathtt{5}: c_2 = \neg\beta$},align=center, base=bottom, node options={dashed}, draw=darkgreen]] %win
				[{$\mathtt{4}: c_1 = \alpha \neq 1$},align=center, base=bottom
				[{$\mathtt{10}: c_2 = \alpha \neq 2$},align=center, base=bottom
				[{$\mathtt{12}: c_3 = \beta$},align=center, base=bottom, node options={dashed}, draw=darkgreen ] % win
				[{$\mathtt{12}: c_3 = \neg\beta$},align=center, base=bottom, node options={dotted,thick}, draw=red]] % loss
				[{$\mathtt{10}: c_2 = \alpha = 2$},align=center, base=bottom
				[{$\mathtt{12}: c_3 = \beta$},align=center, base=bottom, node options={dashed}, draw=darkgreen ] % win
				[{$\mathtt{12}: c_3 = \neg\beta$},align=center, base=bottom, node options={dotted,thick}, draw=red ]]]] % loss
			\end{forest}
		}
		\caption{Execution Tree for the Monty Hall Problem if the car is behind door 1.}
	\end{figure}

	Suppose we were to run traditional symbolic execution on the program in Fig~\ref{fig:montyhall} where the car was randomly chosen to be behind door 1.
	%
	The two inputs, \texttt{choice} and \texttt{door\_switch}, would become the symbolic variables $\alpha$ and $\beta$, respectively.
	%
	The execution tree is shown in Fig.~\ref{fig:montyhall_tree}.
	%
	For each node in the tree the line number of the branch statement and the path condition is given.
	%
	Leaves which are surrounded by a dashed (\tikz{\draw[dashed,line width=1pt] (0,0) -- (0.35,0) }), {\color{darkgreen} green} line represent the contestant winning the car (the function returning \texttt{true}), and those leaves which are surrounded by a dotted (\tikz{\draw[dotted,thick,line width=1pt] (0,0) -- (0.35,0) }), {\color{red} red} line represent the contestant losing the contest (the function returning \texttt{false}).
	%
	Symmetric trees can be made for the cases when the car is behind door 2 and door 3.

	While this tree can tell us that it is \textit{possible} to win (or lose) the car with any choice of initial door and decision to switch, it does not tell us how \textit{often} a contestant will for any initial door and decision to switch.
	%
	In theory, if we knew how likely it was to take one branch over another we could extend this reasoning to determine how often we would hit a winning leaf over a losing leaf.
	%
	\textbf{Our core idea is to represent probabilistic sampling statements also as symbolic variables.}
	%
	For the sake of presentation we refer to these as \textit{probabilistic} symbolic variables as opposed to, what we call, \textit{universal} symbolic variables which range over all possible values.
	%
	So, instead of asking for a random value, we replace the result of the sample with a probabilistic symbolic variable and record the distribution from which the sample is from.
	%
	We then are able to interpret a branch on a probabilistic symbolic variable as a conditioning operation on the originating distribution.

	Now let us return to the Monty Hall problem from Fig.~\ref{fig:montyhall}.
	%
	Let \texttt{car\_door} be represented by the probabilistic symbolic variable, $\delta$ and let \texttt{choice} and \texttt{door\_switch} be $\alpha$ and $\beta$, respectively, as before.
	%
	Instead of branching solely on $\alpha$ and $\beta$, we additionally branch on $\delta$.
	%
	The execution tree is presented in Fig.~\ref{fig:montyhall_tree_prob}.
	%
	Note that branches are omitted from the tree if which direction to traverse can be inferred by the current path condition.

	Since probabilistic symbolic variables originate from a distribution, we can determine the \textit{probability} of taking a certain branch by simply counting how many values from the distribution satisfy the guard condition.
	%
	For example, to figure out the probability of taking the \texttt{true} branch of the \texttt{if} condition on line 4 (denoted by \textbf{T} in Fig.~\ref{montyhall_tree_prob}), it suffices to count how many settings of $\delta$ satisfy $\alpha = \delta$ and divide by the total number of possible assignments to $\delta$ (i.e. 3).
	%
	However, since $\alpha$ is a symbolic variable we can only obtain a probability expression which is in terms of universal symbolic variables.
	%
	Let $[\cdot]$ denote Iverson brackets, where $[Q]=1$ if formula $Q$ is true, and 0 otherwise.
	%
	Then, $\text{\textbf{T}} = ([\alpha = 1] + [\alpha = 2] + [\alpha = 3])/3 = \frac{1}{3}$ as $\delta \in [1,2,3]$ and each setting is equally likely.
	%
	Similarly, $\text{\textbf{F}} = ([\alpha \neq 1] + [\alpha \neq 2] + [\alpha \neq 3])/3 = \frac{2}{3}$.

	\begin{figure}
		\label{fig:montyhall_tree_prob}
		\centering
		{\footnotesize
			\begin{forest}baseline,for tree=draw,
				[{$\mathtt{1}: c_0=\top$},align=center, base=bottom
				[{$\mathtt{4}: c_1 = \alpha = \delta$},align=center, base=bottom, edge label={node [midway,above] {\textbf{T}} }
				[{$\mathtt{5}: c_2 = \beta$},align=center, base=bottom, node options={dotted,thick}, draw=red ] %loss
				[{$\mathtt{5}: c_2 =  \neg\beta$},align=center, base=bottom, node options={dashed}, draw=darkgreen ]] %win
				[{$\mathtt{4}: c_1 = \alpha \neq \delta$},align=center, base=bottom, edge label={node [midway,above] {\textbf{F}} }
				[{$\mathtt{7}: c_2 = \alpha \neq 1 \wedge \delta \neq 1$},align=center, base=bottom, 
				[{$\mathtt{12}: c_3 = \beta$},align=center, base=bottom, node options={dashed}, draw=darkgreen ] %win
				[{$\mathtt{12}: c_3 = \neg\beta$},align=center, base=bottom, node options={dotted,thick}, draw=red ]] %loss
				[{$\mathtt{7}: c_2 = \alpha = 1 \vee \delta = 1$},align=center, base=bottom
				[{$\mathtt{10}: c_3 = \alpha \neq 2 \vee \delta \neq 2$},align=center, base=bottom
				[{$\mathtt{12}: c_4 = \beta$},align=center, base=bottom, node options={dashed}, draw=darkgreen ] %win
				[{$\mathtt{12}: c_4 = \neg\beta$},align=center, base=bottom, node options={dotted,thick}, draw=red ]] %loss
				[{$\mathtt{10}: c_3 = \alpha = 2 \vee \delta = 2$},align=center, base=bottom
				[{$\mathtt{12}: c_4 = \beta$},align=center, base=bottom, node options={dashed}, draw=darkgreen ]%win
				[{$\mathtt{12}: c_4 = \neg\beta$},align=center, base=bottom, node options={dotted,thick}, draw=red ]]]]] %loss
			\end{forest}
		}
		\caption{Execution tree for the Monty Hall Problem with probabilistic symbolic variables.}
	\end{figure}

	With this \textit{probabilistic} execution tree in hand, let's return to the original question we wanted to answer: if the contestant switches doors, does their chances of winning the car exceed $\frac{1}{3}$?
	%
	Note that each path in the tree has a probability associated with it.
	%
	If we want to know the probability of winning if the contestant switches versus not, we can look at solely those paths in the tree which lead to a win.
	%
	To then figure out the probability of winning the car if the contestant switches we then can remove those paths where $\neg\beta$ is true, and then sum up the probabilities of the remaining paths. 
	%
	This results in an expression in terms of $\alpha$ and constants, which we can then evaluate where $\alpha = 1$, $\alpha = 2$, and $\alpha = 3$ and compare the probabilities.
	%
	In the end, we get that regardless the setting of $\alpha$ (i.e. the program variable, \texttt{choice}) the probability of winning the car if the contestant switches doors is exactly $\frac{2}{3}$.
	%
	We formalize this intuition in Section~\ref{sec:query_gen} by expressing this query in first order logic and use automated decision procedures to automate the solving.

	\section{Probabilistic Symbolic Execution Algorithm}
	\label{sec:pse}
	In this section, we present our technique for augmenting traditional symbolic execution to support probabilistic programs with discrete sampling instructions. 
	% 
	We begin with a short review of a traditional symbolic execution algorithm and then discuss how we calculate exact path probabilities.

@@

	% Symbolic Execution Algorithm.
	\begin{algorithm}
		\caption{Probabilistic Symbolic Execution Algorithm}
		\label{alg:symbex}
		\begin{algorithmic}[1]
			\Function{SymbEx}{$P_{prog} : Program$}
			\Let{$\varphi_{paths}$}{[]}, {$Ex_{stack}$} $\leftarrow$ {[]}, $\Delta \leftarrow$ {[]}, $\phi_{paths} \leftarrow \phi$, $P \leftarrow \phi$  \algorithmiccomment{Initialization}
			\Let{$I_{0}$}{\textsc{getStartInstruction}($P_{prog}$)}
			\Let{$S_{0}$}{[$I_{0}$, $\phi_{path}$, $\Delta$, $\phi$]} \algorithmiccomment{Empty Initial State}
			\State {$Ex_{stack}$.\textsc{Push}({$S_{0}$})} \algorithmiccomment{Start with $S_{cur}$ in Execution Stack}
			\While{$Ex_{stack}$ != $\phi$}
			\Let{$S_{cur}$}{$Ex_{stack}$.\textsc{Pop}( )}, $I_{cur} \leftarrow S_{cur}[1]$ \algorithmiccomment{Start State, [$I_{0}$, $\phi_{path}$, $\Delta$]}
			\Switch{\textsc{instType}($I_{cur}$)} \label{line:inst_case}
			\Case{\texttt{make\_pse\_sym}} ($x_{var},  d_{dist}$) \algorithmiccomment{Create \texttt{fresh} Symbolic Variable} \label{line:make_pse}
			\Let{$I_{cur}$}{\textsc{getNextInstruction}($I_{cur}$)}
			\Let{$\phi_{path}, \Delta, P$}{{\textbf{PSESample}}({$x_{var},  d_{dist}$, $\phi_{path}$, $\Delta$, $P$})}
			\Let{$S_{cur}$}{[$I_{cur}$, $\phi_{path}$, $\Delta$, $P$]}
			\State {$Ex_{stack}$.\textsc{Push}({$S_{cur}$})}
			\EndCase
			\Case{v := e} \algorithmiccomment{Assignment Instruction} \label{line:assign}
			\Let{$I_{cur}$}{\textsc{getNextInstruction}($I_{cur}$)}
			\Let{$\phi_{path}, \Delta, P$}{{\textbf{PSEAssignment}}({v, e, $\phi_{path}$, $\Delta$, $P$})}
			\Let{$S_{cur}$}{[$I_{cur}$, $\phi_{path}$, $\Delta$, $P$]}
			\State {$Ex_{stack}$.\textsc{Push}({$S_{cur}$})}
			\EndCase
			\Case{\textbf{if} $(c_{sym})$ \textbf{then} $P_{1}$ \textbf{else} $P_{2}$} \algorithmiccomment{Branch Instruction}
			\Let{$Branch_{expr}$}{{\textbf{PSEBranch}}({$c_{sym}$, $\phi_{path}$, $\Delta$, $P$})} \label{line:branch}
			\State {$S_{true}$}, {$S_{false}$} $\leftarrow$ \textsc{ForkState} ($S_{cur}$, $I_{cur}$) \label{line:fork}
			\Let{$I_{1}$}{\textsc{getStartInstruction}($P_{1}$)} \Let{$I_{2}$} {\textsc{getStartInstruction}($P_{2}$)}	
			\Let{$S_{true}$}{[$I_{1}$, $\phi_{path} \wedge c_{sym}$, $\Delta$, $Branch_{expr}[0]$]} \label{line:true_state}
			\Let{$S_{false}$}{[$I_{2}$, $\phi_{path} \wedge (\neg c_{sym})$, $\Delta$, $Branch_{expr}[1]$]} \label{line:false_state}
			\State {$Ex_{stack}$.\textsc{Push}({$S_{false}$})}
			\State {$Ex_{stack}$.\textsc{Push}({$S_{true}$})} \algorithmiccomment{Start with True State}
			\EndCase
			\Case{\textbf{HALT}} \algorithmiccomment{Terminate Instruction}
			\Let{$\varphi_{paths}$}{$\varphi_{paths}$.\textsc{Append}($\phi_{path}$)}		
			\EndCase
			\EndSwitch
			\EndWhile \\
			\Return{$\varphi_{paths}$}
			\EndFunction
		\end{algorithmic}
	\end{algorithm}

@@

	\subsection{Query Generation}
	\label{sec:query_gen}

	At this point, we have an execution tree where each branch is annotated with the probability of taking the true and false branch.
	% 
	From this tree we construct a set of paths which are represented by a path condition and set of probabilities.
	% 
	If there are $n$ paths, and $\varphi_i$ is the $i^\text{th}$ path condition, and $p_{i_j}$ is the probability of the $j^{\text{th}}$ branch in the $i^{\text{th}}$ path, then we have $\{ (\varphi_1, \{p_{1_1},\ldots,p_{1_{k_1}}\}), \ldots, (\varphi_n, \{p_{n_1},\ldots,p_{n_{k_n}}\})\}$.
	% 
	To calculate the probability of an entire path, or equivalently, the probability of a path condition, $\varphi$, being true, we can simply multiply each of the branch probabilities.
	% 
	Recall that $p_{i_j}$ is the \textit{conditional} probability of the $j^{\text{th}}$ branch condition being true along path $i$.
	% 
	Since $\varphi_i$ is the conjunction of each of the $k_i$ branch conditions, the probability of taking the entire path can be computed using the rule $\Pr[A \wedge B] = \Pr[A \mid B]\Pr[B]$ where $A$ and $B$ are any two events.
	% 
	In other words, since $\varphi_i = c_{i_1} \wedge \cdots \wedge c_{i_{k_i}}$, then
	\begin{align*}
		\Pr[\varphi_i] &= \Pr[c_{i_1} \wedge \cdots \wedge c_{i_{k_i}}]\\
		&= \Pr[c_{i_1} \wedge (c_{i_2} \wedge \cdots \wedge c_{i_{{k_i}-1}} \wedge c_{i_{k_i}})]\\
		&= p_{i_1} \cdot \Pr[c_{i_2} \wedge \cdots \wedge c_{i_{k_i}}]\\
		&\vdots\\
		&= \displaystyle\prod_{j=1}^{k_i} p_{i_j}
	\end{align*}

	So, if we let $p_i = \prod_{j=1}^{k_i} p_{i_j}$, then we can simplify our set of paths to $\Phi = \{ (\varphi_1, p_1), \ldots, (\varphi_n, p_n)\}$.

>>>>

@@

\label{sec:pse}
In this section, we present our technique for augmenting traditional symbolic execution to support probabilistic programs with discrete sampling instructions. 
% 
We begin with a short review of a traditional symbolic execution algorithm and then discuss how we calculate exact path probabilities.

% Symbolic Execution Algorithm.
\begin{algorithm}
	\caption{Probabilistic Symbolic Execution Algorithm}
	\label{alg:symbex}
	\begin{algorithmic}[1]
		\Function{SymbEx}{$Prog : Program$}
		\Let{$\varphi_{paths}$}{[]}, {$Ex_{stack}$} $\leftarrow$ {[]}, $\Delta \leftarrow$ {[]}, $\phi_{paths} \leftarrow \phi$, $P \leftarrow \phi$  \algorithmiccomment{Initialization}
		\Let{$I_{0}$}{\textsc{getStartInstruction}($Prog$)}
		\Let{$S_{0}$}{[$I_{0}$, $\phi_{path}$, $\Delta$, $\phi$]} \algorithmiccomment{Empty Initial State}
		\State {$Ex_{stack}$.\textsc{Push}({$S_{0}$})} \algorithmiccomment{Start with $S_{cur}$ in Execution Stack}
		\While{$Ex_{stack}$ != $\phi$}
		\Let{$S_{cur}$}{$Ex_{stack}$.\textsc{Pop}( )}, $I_{cur} \leftarrow S_{cur}[1]$ \algorithmiccomment{Start State, [$I_{0}$, $\phi_{path}$, $\Delta$]}
		\Switch{\textsc{instType}($I_{cur}$)} \label{line:inst_case}
		\Case{\texttt{make\_pse\_sym}} ($x_{var},  d_{dist}$) \algorithmiccomment{Create \texttt{fresh} Symbolic Variable} \label{line:make_pse}
		\Let{$I_{cur}$}{\textsc{getNextInstruction}($I_{cur}$)}
		\Let{$\phi_{path}, \Delta, P$}{{\textbf{PSESample}}({$x_{var},  d_{dist}$, $\phi_{path}$, $\Delta$, $P$})}
		\Let{$S_{cur}$}{[$I_{cur}$, $\phi_{path}$, $\Delta$, $P$]}
		\State {$Ex_{stack}$.\textsc{Push}({$S_{cur}$})}
		\EndCase
		\Case{v := e} \algorithmiccomment{Assignment Instruction} \label{line:assign}
		\Let{$I_{cur}$}{\textsc{getNextInstruction}($I_{cur}$)}
		\Let{$\phi_{path}, \Delta, P$}{{\textbf{PSEAssignment}}({v, e, $\phi_{path}$, $\Delta$, $P$})}
		\Let{$S_{cur}$}{[$I_{cur}$, $\phi_{path}$, $\Delta$, $P$]}
		\State {$Ex_{stack}$.\textsc{Push}({$S_{cur}$})}
		\EndCase
		\Case{\textbf{if} $(c_{sym})$ \textbf{then} $P_{1}$ \textbf{else} $P_{2}$} \algorithmiccomment{Branch Instruction}
		\Let{$Branch_{expr}$}{{\textbf{PSEBranch}}({$c_{sym}$, $\phi_{path}$, $\Delta$, $P$})} \label{line:branch}
		\Let{$I_{1}$}{\textsc{getStartInstruction}($P_{1}$)} \Let{$I_{2}$} {\textsc{getStartInstruction}($P_{2}$)}	
		\Let{$S_{true}$}{[$I_{1}$, $\phi_{path} \wedge c_{sym}$, $\Delta$, $Branch_{expr}[0]$]} \label{line:true_state}
		\Let{$S_{false}$}{[$I_{2}$, $\phi_{path} \wedge (\neg c_{sym})$, $\Delta$, $Branch_{expr}[1]$]} \label{line:false_state}
		\State {$Ex_{stack}$.\textsc{Push}({$S_{false}$})}
		\State {$Ex_{stack}$.\textsc{Push}({$S_{true}$})} \algorithmiccomment{Start with True State}
		\EndCase
		\Case{\textbf{HALT}} \algorithmiccomment{Terminate Instruction}
		\Let{$\varphi_{paths}$}{$\varphi_{paths}$.\textsc{Append}($\phi_{path}$)}		
		\EndCase
		\EndSwitch
		\EndWhile \\
		\Return{$\varphi_{paths}$}
		\EndFunction
	\end{algorithmic}
\end{algorithm}

@@
